Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-25 ms.tex

\begin{abstract} 
  Social learning is %essential to human survival---it 
  is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of  social learning in populations with cognitively realistic individual learning
  capacities. % We used evolutionary agent-based modeling to track the evolution of social learning versus purely individual learning.
  %how often social, rather than individual,   learning evolved. 
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning. Larger decision set sizes promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

@@

Social learning enhances problem solving when acquiring information from others is more
efficient than learning on one's own~\cite{Laland2004}. However, social learning can also
lead individuals astray if they are copying irrelevant or outdated information.
Theoretical models have helped clarify the circumstances under which social learning
should be evolutionarily favored~\cite{BoydRicherson1985, aoki2014evolution,Kendal2018}, and these models have motivated empirical work that has both validated and refined theoretical predictions concerning social learning behavior in humans
and other species~\cite{galef2005social,McElreath2005,Kendal2018,Allen2019}.  This literature
indicates that social learning is a critical adaptation found across taxa for dealing
with uncertainty. When in doubt, copy others.

Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan societies, and dispersal across a highly variable environments, requiring coarse-grain and plastic behavioral adaptations \cite{levins1962}. However, the term ``uncertainty"
often conflates several different interpretations of that word. For example, different
models define uncertainty in terms of the rate of environmental change, spatial
environmental variation, or the reliability of information acquired from the environment.
Furthermore, in translating mathematical models to verbal predictions, the formal meaning
of terms like uncertainty is often lost \cite{knight1921, lawson1988probability, volz2012}, facilitating such
conflation. %Furthermore, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown~\cite{knight1921,volz2012}. 
As the number of formal models of social learning have expanded, an ever
increasing number of modeling choices~\cite{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models or to consolidate our
understanding of the contexts in which social learning under uncertainty is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

@@

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful mechanism for individul learning. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection set size (the number of
possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for
different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), explained below. In our model, social learners learn from the previous generation and asocial learners do not. All agents engage in individual learning within
their lifespans. The model also assumes a more
cognitively realistic and empirically-motivated learning mechanism, the softmax algorithm, which allows  learners to make full use of the available data acquired by both individual and social learning. 
Our model design thereby affords us a richer and possibly more accurate assessment of tradeoffs between 
social and asocial learning. 
We use this integrated model to ask which kinds of uncertainty are likely to favor social learning and to clarify the logic underlying the results.
%of why this would be the case.  
In the remainder of this introduction we briefly review previous research on both the forms of uncertainty and the learning mechanism used in our model. 
%We then present how our evolutionary agent-based model works. Finally, we explain the logic underlying the result that different forms of uncertainty can either suppress or encourage the evolution of social learning.

@@

\textbf{Temporal environmental variability.} When environments change
(for example, in response to climatic events, migration, or technological paradigm shifts),
strategies that were previously adaptive may no longer be optimal. The difficulty
in predicting either when such shocks will occur or which behaviors will be
adaptive in the resulting environments leads to uncertainty.

Temporal environmental variability is fundamental to most evolutionary models of
social learning. Such fluctuations have been proposed as an important selective
pressure for learned as opposed to genetically fixed behaviors when genetic adaptation cannot keep pace with environmental change 
~\cite{Richerson2000}. On the other hand, environments that change \emph{too} quickly
will select against social learning so that individuals avoid learning outdated--- 
and therefore maladaptive---information~\cite{Feldman1996,
BoydRicherson1985}. This suggests that intermediate levels of temporal variation
are important for the evolution of social learning mechanisms~\cite{aoki2005}.

%Despite the theoretical centrality of t
Temporal environmental variability tends to be modeled in one of 
several ways~\cite{aoki2014evolution}. Most commonly, it is modeled
as an independent probability that the environment changes its state (and its corresponding optimal behavior) before each new generation
~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}, but
% The probabilities of environmental change are usually fixed per generation
%~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, 
it has also been modeled using deterministic cycles, so that environments repeat at regular intervals ~\cite{Feldman1996, aoki2014evolution}.
%Furthermore, t
The consequences of environmental change can range from mild to
catastrophic. In the latter case, a change of environment results in the total
elimination of any adaptive behavior, which must be learned \emph{de novo} for the
new environment~\cite{Rogers1988}. Such catastrophic environmental changes present
a large adaptive challenge as individuals cannot rely on accumulated information
from previous generations. Other models of environmental change introduce less
uncertainty as they fluctuate between two or more environmental states with corresponding
adaptive behaviors. This means that previously maladaptive
behaviors become adaptive when the environment changes
~\cite{perreault2012bayesian}. The chosen mechanism for modelling temporal
change has important theoretical consequences, reviewed in Aoki \& Feldman (2014).~\nocite{aoki2014evolution} Our study design ignores cases where temporal variation is low enough that genetic selection can canalize a behavior, and instead considers only behaviors that can be learned.

@@

A population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. The behavior set is a multi-armed bandit with $B$ arms, that is, agents can pick one of $B$ behaviors with random-valued payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.  
In each generation, exactly one behavior is more likely to pay off than all the rest and is therefore optimal. Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.
%The optimal behavior is static within agent lifespans, but varies in general between generations, representing environmental variability.
At the end of each
generation, agents are selected with replacement to reproduce a new generation of $N$ agents,
with the probability of reproduction biased by net payoffs. Agents who inherit the social
learning trait learn about payoffs from the previous generation, while asocial learners begin
life only knowing the number of behaviors the environment affords, but no knowledge
of behavioral payoffs.
\cm{or would you consider having flat priors over all behaviors knowledge?} 
\mt{I changed this last sentence to express that they have flat priors, in somewhat different
terms.}
%The specifics of the model construction and dynamics are described in detail below.
% \ps{I cut a bunch of stuff that came right here as being redundant or unnecessary.}

@@

\vspace{2em}
\begin{table}[h]
\caption{Environmental parameters used in our simulations, including uncertainty parameters $u$, $B$,
$\pilow$, and $L$. Bold indicates default value tested} %\emph{default value tested}.}
    \label{tab:uncertaintyParameters}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.0in}p{1.25in}} \toprule

@@

All agents perform individual, trial-and-error learning at each time step in
their lifespan.  Learning is guided by softmax search. Softmax search
guides agents to more frequently exploit the most profitable behaviors when the
agent is more certain which is optimal and to explore more frequently when the
agent is unsure.  To do softmax searching, agents track average payoffs acquired
from each behavior they have performed and the number of times they
have performed each behavior. \cm{is the previous sentence still right? I changed it's meaning because it made it sound like tracking average meant keeping track of total number of times it was performed, but that's not inherentely the case. Softmax does BOTH right? to estimate certainty.}  We denote the average payoffs obtained by agent $i$
for behavior $b$ as $\ledger$. We denote the number of times $i$ performed $b$
as $c_{ib}$ (``$c$'' stands for count). 
At each time step, agent $i$ performs behavior $b$ with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}}.
\end{equation}
\noindent
$\beta$ adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$). %~\cite{McElreath2005}. 
We used a default value of $\beta = 10$, but also varied this value in our sensitivity analyses. 
The softmax function used here is a biologically plausible
generalization of behavior selection under uncertainty that enables agents to often
greedily exploit the most lucrative behavior (based on their observations) but also
to sometimes explore alternatives~\cite{Schulz2019,Collins2013,Daw2006,Yechiam2005}.

@@

At the beginning of each generation other than the first, each social learner selects
one member of the previous generation to learn from in a payoff-biased way. 
%, if the child inherited $s_i = 1$. A child agent $c$ is a social learner if its parent $p$ was a social learner, i.e.\ $s_c \leftarrow s_p$, without mutation.
A social learner  
selects this ``teacher" from the previous generation by first selecting $N_T$ prospective teachers at random, then selects its teacher from this set with a probability weighted by each prospective teacher's net payoff 
relative to the other $N_T - 1$ prospective teachers. The social learner then inherits information about the likely payoffs of each behavior from the teacher, whereas asocial learners do not acquire this information. \cm{I think this inheritance has to be explained more. (I see that you do this later, so maybe it can be cut back here)} In this way, the social learner can potentially reduce the amount of exploration needed to both execute and learn about the optimal behavior.

@@

Within each generation, all $N$ agents perform $L$ behaviors sequentially and
independently (Figure~\ref{fig:schematic}B).
Each agent selects one behavior on each step of its lifespan using softmax process
described above, updating the mean payoffs it has observed after each time step. Agent $i$ using behavior $b$ receives a payoff of 1 with probability $\pi_b$. 
%Agents probabilistically recieve a payoff of 0 or 1 depending on whether the bandit paid off. Behavior payoffs are drawn from a Bernoulli distribution with mean $\pi_b$, so $\pi_b$ is also the mean payoff from behavior $b$.  For concreteness we denote this probabilistic payoff $\bandit$. 
After performing behavior $b$, agent $i$ updates the
corresponding \emph{behavior count} by 1, $c_{ib} \leftarrow c_{ib} + 1$, and then
the expected payoffs calculated for that behavior are updated with
exponentially-weighted averaging
\begin{equation}
  \ledger \leftarrow \ledger + \frac{\bandit - \ledger}{c_{ib}},
\end{equation}
\noindent
where $\bandit$ is the actual payoff received by agent $i$ for behavior $b$ on the given time step.

\paragraph{Intergenerational dynamics}
Between generations, agents first reproduce via asexual haploid reproduction, which determines the transmission of the social learning trait. Social learners then learn from the
previous generation, after which all agents from the
previous generation die off (Figure~\ref{fig:schematic}C). More specifically, this all happens as follows. 
$N$ reproducers are selected with replacement over $N$ independent draws, 
biased by performance:
\begin{equation}
  \Pr(\text{Agent $i$ is chosen to reproduce in one draw}) = \frac{\pi_i}{\sum_{i=1}^N \pi_i}.
\end{equation}
\noindent
A child inherits its parent's social learning trait $s_i$ without mutation.
A social learner child with $s_i = 1$ learns from a teacher from its parent's
generation, including possibly their parent (i.e., both vertical and oblique learning are possible). 
A child chooses its teacher by first randomly selecting $N_T$ prospective
teachers from the population, then selecting the one with
highest net payoffs $\pi_i$ as their teacher, with ties broken randomly \cm{this seems to contradict info under 2.2.2 Social learning section where there seems to be a probabilistic element weighted by relative payoff of teachers. I think it doesn't have to be described in both places.}. By first subsetting $N_T$ prospective teachers we represent the fact that access to the entire population is not generally guaranteed, though this type of algorithm yields qualitatively similar results to ones that do not first produce subsamples \cite{smaldino2019open}. 
Social learners adopt their chosen teacher's expected payoffs observations $\ledger$, \ps{I feel
like there should be some indication that the ledger is a vector rather than a scalar. Maybe put
it in bold or capitalize the $\Pi$? Using the overbar makes it look like a single average.}
\mt{The subscript $b$ makes it clear it's a vector over behaviors. Maybe we lose the bar,
or I may change the notation to at least approximately follow Q-learning notation.}
and set all $c_{ib} = 1$ so expected payoffs remain within $[0, 1]$. Setting $c_{ib}$ to 1 reflects the fact that the social learning is treating the teacher's information as a single observation with the associated uncertainty.   Asocial
learners ($s_i = 0$) are initialized with $c_{ib} = 0$ and $\ledger = 0$ 
(Figure~\ref{fig:schematic}C). \ps{How to account for division by zero?} \mt{Count increases
before average are taken, which I think is clear enough in paragraph \textbf{Within-generation
dynamics}.} The
newly initialized population then repeats the within-generation dynamics.

@@

\begin{figure}
  \caption{Agents are randomly initialized as social learners or not, with their
  payoff observations all initialized to zero (A). Then agents begin selecting
and performing behaviors and accumulating payoffs, which goes on for $L$
timesteps (B). After $L$ time steps, agents are selected to reproduce,
social learner children learn from a member of their parent's generation, and
the previous generation dies off (C). The simulation stops if children are all
social or asocial learners (i.e.\ the system reaches fixation), 
otherwise repeat another generation and evolution (3).}
  \label{fig:schematic}
  \centering
    \includegraphics[width=\textwidth]{Figures/IntraInterGenerationalDynamics.pdf}
\end{figure}

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values (Table~\ref{tab:modelParameters}) and observed how frequently social learning evolved across
1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
We observed three outcome measures, with $\meansl$ as the primary measure: the average value of $s_i$ over all agents and trials. To support our conclusions
based on $\meansl$, we also measured $\meanpi / L$, the average net  payoffs at simulation end across agents and trials, normalized by lifespan; 
and $\meanG$, the mean
number of generations to fixation (see Table~\ref{tab:outcomeVariables}). We analyze outcomes by plotting $\meansl$ on the y-axis and environmental variation is on x-axis since we
theoretically expect that $\meansl$ will decrease monotonically from 1 to 0 as $u$
increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

\begin{table}[h]
    \caption{Outcome variables.}
    \label{tab:outcomeVariables}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.25in}p{0.85in}} \toprule

@@

        $\meansl$ & Mean social learning prevalence over agents and trials
                  & $\in [0.0, 1.0]$ \\

        $\meanpi / L$ & Mean payoffs accumulated in a generation normalized by
        lifespan & $\in [0.0, 1.0]$ \\

        $\meanG$ & Mean number of generations to convergence & 20k / 8 max. \\
        \bottomrule
    \end{tabular}
\end{table}

>>>>

\begin{abstract} 
  Social learning %essential to human survival---it 
  is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of social learning in populations with cognitively-realistic individual-learning
  capacities. % We used evolutionary agent-based modeling to track the evolution of social learning versus purely individual learning.
  %how often social, rather than individual,   learning evolved. 
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning, whereas larger decision-set size promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

@@

Social learning enhances problem solving when acquiring information from others is more
efficient than learning on one's own~\cite{Laland2004}. However, social learning can also
lead individuals astray if they are copying irrelevant, misleading, or outdated information.
Theoretical models have helped clarify the circumstances under which social learning
should be evolutionarily favored~\cite{BoydRicherson1985, aoki2014evolution,Kendal2018}, and these models have motivated empirical work that has both validated and refined theoretical predictions concerning the use of social learning in humans
and other species~\cite{galef2005social,McElreath2005,Kendal2018,Allen2019}.  This literature
indicates that social learning is a critical adaptation found across taxa for dealing
with uncertainty. When in doubt, copy others.

Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan societies, and dispersal across a highly variable environments, requiring coarse-grain and plastic behavioral adaptations \cite{levins1962}. However, the term ``uncertainty''
often conflates several different interpretations of that word. For example, different
models define uncertainty in terms of the rate of environmental change, spatial
environmental variation, or the reliability of information acquired from the environment.
Furthermore, in translating mathematical models to verbal predictions, the formal meaning
of terms like uncertainty is often lost \cite{knight1921, lawson1988probability, volz2012}, facilitating such
conflation. %Furthermore, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown~\cite{knight1921,volz2012}. 
As the number of formal models of social learning have expanded, an increasing number of modeling choices~\cite{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models or to consolidate our
understanding of the contexts in which social learning under uncertainty is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

@@

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several forms of uncertainty and endows agents
with a relatively powerful mechanism for individual learning. We model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection set size (the number of
possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for
different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), explained below. In our model, social learners learn from the previous generation and asocial learners do not. All agents engage in individual learning within
their lifespans. The model also assumes a more
cognitively-realistic and empirically-motivated learning mechanism, the softmax algorithm, which allows learners to make full use of the available data acquired by both individual and social learning. 
Our model design thereby affords us a richer and possibly more accurate assessment of trade-offs between 
social and asocial learning. 
We use this integrated model to ask which kinds of uncertainty are likely to favor social learning and to clarify the logic underlying the results.
%of why this would be the case.  
In the remainder of this introduction we briefly review previous research on both the forms of uncertainty and the learning mechanism used in our model. 
%We then present how our evolutionary agent-based model works. Finally, we explain the logic underlying the result that different forms of uncertainty can either suppress or encourage the evolution of social learning.

@@

\textbf{Temporal environmental variability.} When environmental variability is nonstationary (for example, because of climatic events, migration, or technological paradigm shifts), strategies that were previously adaptive may no longer be optimal. The difficulty in predicting either when such shocks will occur or which behaviors will be adaptive in the resulting environments leads to uncertainty.

Temporal environmental variability is fundamental to most evolutionary models of
social learning. Such fluctuations have been proposed as an important selective
pressure for learned as opposed to genetically fixed behaviors when genetic adaptation cannot keep pace with environmental change~\cite{Richerson2000}. On the other hand, environments that change \emph{too} quickly will select against social learning so that individuals avoid learning outdated---and therefore maladaptive---information~\cite{Feldman1996,
BoydRicherson1985}. This suggests that intermediate levels of temporal variation
are important for the evolution of social learning mechanisms~\cite{aoki2005}.

%Despite the theoretical centrality of t
Temporal environmental variability tends to be modeled in one of 
several ways~\cite{aoki2014evolution}. Most commonly, it is modeled
as an independent probability that the environment changes its state (and its corresponding optimal behavior) before each new generation
~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}, but
% The probabilities of environmental change are usually fixed per generation
%~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, 
it has also been modeled using deterministic cycles, so that environments repeat at regular intervals ~\cite{Feldman1996, aoki2014evolution}.
%Furthermore, t
The consequences of environmental change can range from mild to
catastrophic. In the latter case, a change of environment results in the total
elimination of any adaptive behavior, which must be learned \emph{de novo} for the
new environment~\cite{Rogers1988}. Such catastrophic environmental changes present
a large adaptive challenge as individuals cannot rely on accumulated information
from previous generations. Other models of environmental change introduce less
uncertainty as they fluctuate between two or more environmental states with corresponding adaptive behaviors. This means that previously maladaptive
behaviors become adaptive when the environment changes~\cite{perreault2012bayesian}. The chosen mechanism for modelling temporal
change has important theoretical consequences, reviewed in Aoki \& Feldman (2014).~\nocite{aoki2014evolution} Our study design ignores cases where temporal variation is low enough that genetic selection can canalize a behavior, and instead considers only behaviors that can be learned.

@@

A population of $N$ individuals each must decide which of $B$ behaviors to perform at
each time step within a generation consisting of $L$ time steps. The behavior set is a multi-armed bandit with $B$ arms, that is, agents can pick one of $B$ behaviors with random-valued payoffs~\cite{SuttonBartoBook,McElreath2005,Steyvers2009,Rendell2010,Schulz2019}.  
In each generation, exactly one behavior is more likely to pay off than all the rest and is therefore optimal. Agents
optimize their net payoffs over their lifespan by quickly figuring out which
behavior is optimal and performing it as often as possible within their lifespan.
%The optimal behavior is static within agent lifespans, but varies in general between generations, representing environmental variability.
At the end of each
generation, agents are selected with replacement to reproduce a new generation of $N$ agents,
with the probability of reproduction biased by net payoffs. Agents who inherit the social
learning trait learn about payoffs from the previous generation, while asocial learners begin
life only knowing the number of behaviors the environment affords, but no knowledge
of behavioral payoffs.
\cm{or would you consider having flat priors over all behaviors knowledge?} 
\mt{I changed this last sentence to express that they have flat priors, in different
terms.}
%The specifics of the model construction and dynamics are described in detail below.
% \ps{I cut a bunch of stuff that came right here as being redundant or unnecessary.}

@@

\vspace{2em}
\begin{table}[h]
\caption{Environmental Parameters. These include our four main uncertainty parameters under investigation; $u$, $B$,
$\pilow$, and $L$. Bold indicates default value tested} %\emph{default value tested}.}
    \label{tab:uncertaintyParameters}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.0in}p{1.25in}} \toprule

@@

All agents perform individual, trial-and-error learning at each time step in
their lifespan.  Learning is guided by softmax search. Softmax search
guides agents to more frequently exploit the most profitable behaviors when the
agent is more certain which is optimal and to explore more frequently when the
agent is unsure.  To do softmax searching, agents track average payoffs acquired
from each behavior they have performed, which requires knowing the number of times they
have performed each behavior. The probability that the agent will perform a behavior is a function of the agent's beliefs about a behavior's average payoff in that timestep and a fixed parameter that influences the amount of exploratory behavior (more details below). The softmax function used here is a biologically plausible generalization of behavior selection under uncertainty that enables agents to often greedily exploit the most lucrative behavior (based on their observations) but also to sometimes explore alternatives~\cite{Schulz2019,Collins2013,Daw2006,Yechiam2005}.

@@

At the beginning of each generation other than the first, each social learner selects
one member of the previous generation to learn from in a payoff-biased way. 
%, if the child inherited $s_i = 1$. A child agent $c$ is a social learner if its parent $p$ was a social learner, i.e.\ $s_c \leftarrow s_p$, without mutation.
A social learner selects this ``teacher" from the previous generation by first choosing a random subset of the population and choosing among these with a probability weighted by each model's payoff (more details below).  The social learner then inherits information about the likely payoffs of each behavior from that one teacher, whereas asocial learners do not acquire this information. In this way, the social learner can potentially reduce the amount of exploration needed to both execute and learn about the optimal behavior.

@@

Within each generation, all $N$ agents perform $L$ behaviors sequentially and
independently (Figure~\ref{fig:schematic}B).
Each agent selects one behavior on each step of its lifespan using softmax process that integrates previous knowledge. We denote the average payoffs obtained by agent $i$
for behavior $b$ as $\ledger$. We denote the number of times $i$ performed $b$
as $c_{ib}$ (``$c$'' stands for count). 
At each time step, agent $i$ performs behavior $b$ with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}}.
\end{equation}
\noindent
$\beta$ adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$). %~\cite{McElreath2005}. 
We used a default value of $\beta = 10$, but also varied this value in our sensitivity analyses.

Agent $i$ using behavior $b$ receives a payoff of 1 with probability $\pi_b$. 
%Agents probabilistically recieve a payoff of 0 or 1 depending on whether the bandit paid off. Behavior payoffs are drawn from a Bernoulli distribution with mean $\pi_b$, so $\pi_b$ is also the mean payoff from behavior $b$.  For concreteness we denote this probabilistic payoff $\bandit$. 
After performing behavior $b$, agent $i$ updates the
corresponding \emph{behavior count} by 1, $c_{ib} \leftarrow c_{ib} + 1$, and then
the expected payoffs calculated for that behavior are updated with
exponentially-weighted averaging
\begin{equation}
  \ledger \leftarrow \ledger + \frac{\bandit - \ledger}{c_{ib}},
\end{equation}
\noindent
where $\bandit$ is the actual payoff received by agent $i$ for behavior $b$ on the given time step.

\paragraph{Intergenerational dynamics}
Between generations, agents first reproduce via asexual haploid reproduction, which determines the transmission of the social learning trait. Social learners then learn from the
previous generation, after which all agents from the
previous generation die off (Figure~\ref{fig:schematic}C). More specifically, this all happens as follows. 
$N$ reproducers are selected with replacement over $N$ independent draws, 
biased by performance:
\begin{equation}
  \Pr(\text{Agent $i$ is chosen to reproduce in one draw}) = \frac{\pi_i}{\sum_{i=1}^N \pi_i}.
\end{equation}
\noindent
A child inherits its parent's social learning trait $s_i$ without mutation.
A social learner child with $s_i = 1$ learns from a teacher from its parent's
generation, including possibly their parent (i.e., both vertical and oblique learning are possible). 
A child chooses its teacher by first randomly selecting $N_T$ prospective
teachers from the population, then selecting one from this set with a probability weighted by each prospective teacher's net payoff, $\pi_i$,  
relative to the other $N_T - 1$ prospective teachers, with ties broken randomly. By first subsetting $N_T$ prospective teachers we represent the fact that access to the entire population is not generally guaranteed, though this type of algorithm yields qualitatively similar results to ones that do not first produce subsamples \cite{smaldino2019open}. 
Social learners adopt their chosen teacher's expected payoffs observations $\ledger$, \ps{I feel
like there should be some indication that the ledger is a vector rather than a scalar. Maybe put
it in bold or capitalize the $\Pi$? Using the overbar makes it look like a single average.}
\mt{The subscript $b$ makes it clear it's a vector over behaviors. Maybe we lose the bar,
or I may change the notation to at least approximately follow Q-learning notation.}
and set all $c_{ib} = 1$ so expected payoffs remain within $[0, 1]$. Setting $c_{ib}$ to 1 reflects the fact that the social learning is treating the teacher's information as a single observation with the associated uncertainty.   Asocial
learners ($s_i = 0$) are initialized with $c_{ib} = 0$ and $\ledger = 0$ 
(Figure~\ref{fig:schematic}C). \ps{How to account for division by zero?} \mt{Count increases
before average are taken, which I think is clear enough in paragraph \textbf{Within-generation
dynamics}.} The
newly initialized population then repeats the within-generation dynamics.

@@

\begin{figure}
  \caption{Agents are randomly initialized as social learners or not, with their
  payoff observations all initialized to zero (A). Then agents begin selecting
and performing behaviors and accumulating payoffs, which goes on for $L$
timesteps (B). After $L$ time steps, agents are selected to reproduce,
social learner children learn from a member of their parent's generation, and
the previous generation dies off (C). The simulation stops if children are all
social or asocial learners (i.e.\ the system reaches fixation), 
otherwise repeat another generation and evolution (3).} \cm{What does (3) denote at end of Fig 1 caption? Does having only green agents in Fig1B give the impression that only social learners are learning within lifetime? Could the agent be half green/purple? Is it weird that in Fig1C a teacher with relatively low payoffs is being selected? Could be fixed by having the dotted line go from rich purple women's head to green baby's. }
  \label{fig:schematic}
  \centering
    \includegraphics[width=\textwidth]{Figures/IntraInterGenerationalDynamics.pdf}
\end{figure}

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values (Table~\ref{tab:modelParameters}) and observed how frequently social learning evolved across
1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
We observed three outcome measures, with $\meansl$ as the primary measure: the average value of $s_i$ over all agents and trials. To support our conclusions
about $\meansl$, we also measured $\meanpi / L$, the average net  payoffs at the end of the simulation across agents and trials, normalized by lifespan; 
and $\meanG$, the mean
number of generations to fixation (see Table~\ref{tab:outcomeVariables}). We analyze outcomes by plotting $\meansl$ on the y-axis and environmental variability is on x-axis since we
theoretically expect that $\meansl$ will decrease monotonically from 1 to 0 as $u$
increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

\begin{table}[h]
    \caption{Outcome Variables. All averages are computed across trials at the end of the last generation.}
    \label{tab:outcomeVariables}
    \centering %\hspace{-3em}
    \begin{tabular}{cp{4.25in}p{0.85in}} \toprule

@@

        $\meansl$ & Mean social learning prevalence over agents
                  & $\in [0.0, 1.0]$ \\

        $\meanpi / L$ & Mean payoffs accumulated across agents, normalized by
        lifespan & $\in [0.0, 1.0]$ \\

        $\meanG$ & Mean number of generations to convergence & 100k / 8 max. \\
        \bottomrule
    \end{tabular}
\end{table}
